# Activity Monitoring Data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
df <- read.csv("activity.csv", sep=",", header = TRUE)
df$date <- as.Date(df$date)
summary(df)
str(df)
```

## 1.Calculating Mean steps taken per day

```{r}
library(ggplot2)
s <- aggregate(steps ~ date, data=df, sum, rm.na=TRUE)
p <- ggplot(s, aes(x=steps)) + 
  geom_histogram(color="black", fill="white")
p
```

Here, we can calculate mean and median steps of each day

```{r}
mean(s$steps)
median(s$steps)
```
Both are almost the same.

## 2. Average Daily Activity Pattern

```{r}
avgSteps <- aggregate(steps ~ interval, data=df, mean, rm.na=TRUE)
q <- ggplot(data = avgSteps, aes(x=interval, y=steps)) +
  geom_line()
q
avgSteps[(which.max(avgSteps$steps)),]
```

**835th** Interval has the maximum average number of steps.

## 3. Imputing Missing Values

```{r}
# Creating a new dataset for imputing NA's
df_imput <- df

# Function for mapping intervals 
fillingNA <- function(interval){
  avgSteps[avgSteps$interval==interval,"steps"]
}
  
for(i in 1:nrow(df_imput)){
  if(is.na(df_imput[i,"steps"])){
    df_imput[i,"steps"] <- fillingNA(df_imput[i,"interval"])
  }
}
head(df_imput)

# Plotting the new dataset with mean steps for each day
s1 <- aggregate(steps ~ date, data=df_imput, sum, rm.na=TRUE)
r <- ggplot(data = s1, aes(x=steps)) +
  geom_histogram(color="black", fill="white")
r

# Mean and Median for the imputed dataset
mean(s1$steps)
median(s1$steps)
```
Only value that differ from the missing dataset is the **median**. **Mean** value remains the same.

## 4. Pattern in weekdays and weekends

```{r}
## Splitting days as Weekday and Weekend
df_imput$weekday <- ifelse(as.POSIXlt(as.Date(df_imput$date))$wday%%6 == 0, "weekend","weekday")
df_imput$weekday <- as.factor(df_imput$weekday)
str(df_imput)
```
```{r}
## Plotting activity weekdays vs weekends
avgSteps1 <- aggregate(steps ~ interval + weekday, data=df_imput, mean)

t <- ggplot(data = avgSteps1, aes(x=interval, y=steps)) +
  geom_line() + facet_grid(weekday ~ .)
t
```

During weekdays, the average number of steps has reached upto 225 whereas durng weekends it doesn't exceed 175.
